---
title: Hands-on training session 2
subtitle: Hui-Walter models for diagnostic test evaluation
date: "`r Sys.Date()`"
fontsize: 12pt
author:
  - Matt Denwood
  - Giles Innocent
theme: metropolis
aspectratio: 43
colortheme: seahorse
header-includes: 
  - \input{preamble}
fig_caption: true
classoption: compress, c
output:
    beamer_presentation:
        pandoc_args: ["-t", "beamer"]
        slide_level: 2
    html_document: default
---

```{r rendering, eval=FALSE, include=FALSE}
# To render this as both html and PDF (beamer) slides run:
rmarkdown::render('Adv2_HuiWalter.Rmd', 'all')
# Or just for html:
rmarkdown::render('Adv2_HuiWalter.Rmd', 'html_document')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

## Overview

Date/time:

  - 19th February 2020
  - 16.00 - 17.00

Teachers:

  - Matt Denwood (presenter)
  - Giles Innocent

## Recap

Important points from session 1




# Session 2a:  Hui-Walter models for 2 tests and 1 population

## Model specification

## Exercise


# Session 2b:  Hui-Walter models for 2 tests and N populations

## Model specification

Random effect of population vs independent intercepts

## Exercise



## Session 2 (1 hour: Multiple tests)
### 2 Tests, 1 Population
What do we mean by "conditionally independent?"
Df in the model and in the data
Use of informative priors
```{r 2 test 1 population, informative priors}
# R code simulating data
# Jags/R code analysing data
# R code to produce appropriate output
```
### Experiments
What happens as you reduce the information in the priors?
### 2 tests, 2+ Populations
Hui Walter model
```{r 2 test 2 poulation}
# R code simulating data
# Jags/R code analysing data
# R code to produce appropriate output
```
### Experiment
Try setting both prevalences very close to 0.5
Noting that there are 2 solutions and model can jump between solutions. Restricting the solution to teh preferred one by forcing Se+Sp>1
