---
title: Hands-on training session 3
subtitle: Hui-Walter models with more than two diagnostic tests
date: "`r Sys.Date()`"
author:
  - Matt Denwood
  - Giles Innocent
  - Sonja Hartnack
theme: metropolis
aspectratio: 43
colortheme: seahorse
header-includes: 
  - \input{preamble}
params:
  presentation: TRUE
output:
  beamer_presentation:
      pandoc_args: ["-t", "beamer"]
      slide_level: 2
  html_document: default
---

```{r rendering, eval=FALSE, include=FALSE}
# To render this as PDF (beamer) slides run:
rmarkdown::render('Adv3_MultipleTests.Rmd', 'beamer_presentation', params=list(presentation=TRUE))
# And for html:
rmarkdown::render('Adv3_MultipleTests.Rmd', 'html_document', params=list(presentation=FALSE))
```

```{r setup, include=FALSE}
# Reduce the width of R code output for PDF only:
if(params$presentation) options(width=60)
knitr::opts_chunk$set(echo = TRUE)

library('runjags')
runjags.options(silent.jags=TRUE)
```


# Introduction

## Overview

Date/time:

  - 20th February 2020
  - 14.00 - 15.30

Teachers:

  - Matt Denwood (presenter)
  - Giles Innocent
  - Sonja Hartnack


## Recap

Important points from sessions 1 and 2




# Session 3a:  Hui-Walter models for multiple tests with conditional indepdendence

## What exactly is our latent class?

What do we mean by "conditionally independent?"

Example:  three antibody tests

The latent status is actually 'producing antibodies' not 'diseased'

We're actually pulling **something** out of a hat, and deciding to call it a rabbit

## Model specification

If doing this manually, take **extreme** care with multinomial tabulation

Or use autohuiwalter!

  * This will also deal gracefully with missing data in one or more test results


## Simulating data

Show how to simulate data for arbitrary numbers of test results


## Exercise

Simulate data from 3 or 4 tests and analyse


## Optional Exercise

Make some data missing for one or more tests and re-generate the model

  * Can you see what has changed in the code?


# Session 3b:  Hui-Walter models for multiple tests with conditional depdendence

## Branching of processes leading to test results

Example:  two antibody tests and one antigen test

Or three antibody tests where one has a different target to the others

## Model specification

Introduce correlations between se and sp

Use autohuiwalter!

## Simulating data

Show how to simulate data using an explicit branching process, and correct Se/Sp estimates

[Steal code from ABME course:  ABME_hui_walter_complete.R]


## Exercise

Simulate data with a dependence between 2 tests

Model assuming conditional independence biases the estimates

Model with conditional depdendence has bigger CI but unbiased


# Session 3c:  Model selection

## Motivation

[Planning for this session to be a general discussion between all instructors and students, as I am not entirely sure what to recommend in terms of model selection - except that I dislike DIC!!!]

## Background to DIC

[Some theory slides stolen from ABME course:  ABME_Model selection.pptx]



## Other methods

DIC works fine for hierarchical normal models

Bayes factors work well if you can count them

WAIC works better for a wide range of models
	
	* Probably won't work for Hui-Walter though due to lack of independent data
	* Could be useful if using the GLM version (untested!)

Models tend to be sensitive to priors

Simulating data and testing that your model recovers the parameters is a good idea

## Calculating DIC

Add dic and ped to the monitors in runjags

But don't trust the results

Also bear in mind you can't parallelise

## Calculating WAIC

Currently a pain - steal code from ABME course:  ABME_waic_example.R


## Future Updates

Model criticism will get better in JAGS 5, and the next update of runjags

Installing development version of runjags:

```{r}
# Put on drat server and supply code here
```

WAIC is also calculable from Stan models (easily?)

## Discussion and free practical time

What would be useful to add to the autohuiwalter function?

  * Modify so it allows Se/Sp priors to be defined as matrices?
  * And correlations on/off as matrices?
