---
title: Hands-on training session 3
subtitle: Hui-Walter models with more than two diagnostic tests
date: "`r Sys.Date()`"
fontsize: 12pt
author:
  - Matt Denwood
  - Giles Innocent
  - Sonja Hartnack
theme: metropolis
aspectratio: 43
colortheme: seahorse
header-includes: 
  - \input{preamble}
fig_caption: true
classoption: compress, c
output:
    beamer_presentation:
        pandoc_args: ["-t", "beamer"]
        slide_level: 2
    html_document: default
---

```{r rendering, eval=FALSE, include=FALSE}
# To render this as both html and PDF (beamer) slides run:
rmarkdown::render('Adv3_MultipleTests.Rmd', 'all')
# Or just for html:
rmarkdown::render('Adv3_MultipleTests.Rmd', 'html_document')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

## Overview

Date/time:

  - 20th February 2020
  - 14.00 - 15.30

Teachers:

  - Matt Denwood (presenter)
  - Giles Innocent
  - Sonja Hartnack


## Recap

Important points from sessions 1 and 2




# Session 3a:  Hui-Walter models for multiple tests with conditional indepdendence

## Model specification

## Exercise


# Session 3b:  Hui-Walter models for multiple tests with conditional depdendence

## Model specification

Being careful!

Give them my function

## Exercise


# Session 3c:  Model selection

## Methods of selecting models

DIC works fine for hierarchical normal models

Bayes factors work well if you can count them

WAIC works better for a wide range of models, although perhaps not these unless specifying the model in a different way (loop over individuals)

Models tend to be sensitive to priors

Simulating data and testing that your model recovers the parameters is a good idea

## Exercise



## Session 3 (1 hour 30 mins: Developing skills)
### 3 tests 1 population
Conditional independence
```{r 3 test 1 pop}
# R code simulating data
# Jags/R code analysing data
# R code to produce appropriate output
```
### 3 tests 1 population: conditional dependence
How do we code conditional dependence?
```{r 3 test 1 pop dependence}
# R code simulating data
# Jags/R code analysing data
# R code to produce appropriate output
```
### Model Selection
Not my problem: that's for you and Sonja, but presumably you'll talk about AIC/BIC/DIC and why (not) to use them and 
then Bayes factors

They also can be linked to R code, which allows us to produce a wide variety of post-analysis summaries and graphical representations.

```{r 1 population, 1 test}
# R code simulating data
# Parameters
Prev1 <- 0.8  # prevalence
Se1 <- 0.8    # test sensitivity
Sp1 <- 0.95   # test specificity
n.obs <- 50   # Number of individuals tested (number of observations)
n.burnin <- 2000  # number of burn-in iterations for MCMC
n.sample <- 2000  # number of samples to take after burn-in
# simulation
true.positive <- rbinom(1, n.obs, Prev1)
test.positive <- rbinom(1, true.positive, Se1)
test.positive <- test.positive + rbinom(1, n.obs-true.positive, 1-Sp1)
# Jags/R code analysing data
```

