---
title: Hands-on training session 4
subtitle: BLCMs for Se & Sp estimation with covariates
date: "`r Sys.Date()`"
author:
  - Sonja Hartnack
  - Eleftherios Meletis
theme: metropolis
aspectratio: 43
colortheme: seahorse
header-includes: 
  - \input{preamble}
params:
  presentation: TRUE
output:
  beamer_presentation:
      pandoc_args: ["-t", "beamer"]
      slide_level: 2
  html_document: default
---

```{r rendering, eval=FALSE, include=FALSE}
# To render this as PDF (beamer) slides run:
rmarkdown::render('Adv4_Covariates.Rmd', 'beamer_presentation', params=list(presentation=TRUE))
# And for html:
rmarkdown::render('Adv4_Covariates.Rmd', 'html_document', params=list(presentation=FALSE))
```

```{r setup, include=FALSE}
# Reduce the width of R code output for PDF only:
if(params$presentation) options(width=60)
knitr::opts_chunk$set(echo = TRUE)

library('runjags')
runjags.options(silent.jags=TRUE)
```


# Session 4:  BLCMs for Se & Sp estimation with covariates

## Overview

Date/time:

  - 20th February 2020
  - 16.00 - 17.00

Teachers:

  - Sonja Hartnack (presenter)
  - Eleftherios Meletis
  
## Recap

We looked at a GLM-type formulation of this model yesterday:

```{r include=FALSE}
glmhw_definition <- "model{

  for(i in 1:N){
    Status ~ dcat(prob[i, ])
  
	  prob[i,1] <- (prev[i] * ((1-se[1])*(1-se[2]))) + 
	              ((1-prev[i]) * ((sp[1])*(sp[2])))
	  prob[i,2] <- (prev[i] * ((se[1])*(1-se[2]))) + 
	              ((1-prev[i]) * ((1-sp[1])*(sp[2])))
	  prob[i,3] <- (prev[i] * ((1-se[1])*(se[2]))) + 
	              ((1-prev[i]) * ((sp[1])*(1-sp[2])))
	  prob[i,4] <- (prev[i] * ((se[1])*(se[2]))) + 
	              ((1-prev[i]) * ((1-sp[1])*(1-sp[2])))
	  
	  logit(prev[i]) <- intercept + population_effect[Population[i]]
  }

  intercept ~ dnorm(0, 0.33)
  population_effect[1] <- 0
  for(p in 2:Pops){
    population_effect[p] ~ dnorm(0, 0.1)
  }
  se[1] ~ dbeta(HPSe[1,1], HPSe[1,2])T(1-sp[1], )
  sp[1] ~ dbeta(HPSp[1,1], HPSp[1,2])
  se[2] ~ dbeta(HPSe[2,1], HPSe[2,2])T(1-sp[2], )
  sp[2] ~ dbeta(HPSp[2,1], HPSp[2,2])

  #data# Status, N, Population, Pops, HPSe, HPSp
  #monitor# intercept, population_effect, se, sp
  #inits# intercept, population_effect, se, sp
}
"
cat(glmhw_definition, file='glm_hw.bug')
```


```{r comment='', echo=FALSE}
cat(glmhw_definition, sep='\n')
```

---

We can think of "Population" as a covariate, and add others.  For example:

```{r echo=FALSE}
set.seed(2020-02-12)
```

```{r}
se1 <- 0.9
sp1 <- 0.95
sp2 <- 0.99
se2 <- 0.8

int <- -1.5
pop_eff <- c(0, 1.5, 3)
sex_eff <- c(0, 0.2)
age_eff <- 0.1
N <- 300

simdata <- data.frame(
  Population = sample(seq_along(pop_eff), N, replace=TRUE),
  Sex = sample(seq_along(sex_eff), N, replace=TRUE),
  Age = runif(N, -3, 3)
  )
simdata$logitprob <- with(simdata, int + pop_eff[Population] + sex_eff[Sex] + age_eff*Age)

simdata$status <- rbinom(N, 1, plogis(simdata$logitprob))
simdata$Test1 <- rbinom(N, 1, (simdata$status * se1) + ((1-simdata$status) * (1-sp1)))
simdata$Test2 <- rbinom(N, 1, (simdata$status * se2) + ((1-simdata$status) * (1-sp2)))
  
head(simdata)
```


---

We can cheat a bit and use template.jags to help set up part of the model:

```{r}
simdata$Population <- factor(simdata$Population, levels=seq_along(pop_eff), labels=paste0('Pop_', seq_along(pop_eff)))
simdata$Sex <- factor(simdata$Sex, levels=seq_along(sex_eff), labels=c('female','male'))

runjags::template.jags(status ~ Population + Sex + Age, data=simdata, file='glmtemplate.bug', family='binomial', effect.prior = 'dnorm(0, 0.1)', write.data=FALSE)
```

The lines you need from this auto-generated model are #17:

```{r echo=FALSE}
cat(readLines('glmtemplate.bug')[17], sep='\n')
```

[And change `regression_prob´ to `prev`]

---

And also lines 21:27 - changing priors as needed:

```{r echo=FALSE}
cat(readLines('glmtemplate.bug')[21:27], sep='\n')
```

---

Line 33 may also be helpful (we don't need deviance, dic or resid.sum.sq):

```{r echo=FALSE}
cat(readLines('glmtemplate.bug')[33], sep='\n')
```

---

And lines 39-57 for initial values:

```{r echo=FALSE}
cat(readLines('glmtemplate.bug')[45:57], sep='\n')
```

---

Combine the two models, add initial values for se and sp, and also add a `#data# N, Status, Age, Population, Sex´ block, and modify the priors for test1 se/sp.

```{r echo=FALSE}
if(!params$presentation) cat(readLines('HW_GLM_coefs.bug'), sep='\n')
```


---

To run the model:

```{r}
N <- nrow(simdata)
simdata$Status <- with(simdata, factor(interaction(Test1, Test2), levels=c('0.0','1.0','0.1','1.1')))
```

```{r, eval=FALSE}
results <- run.jags('HW_GLM_coefs.bug', data=simdata)
results
```

---

```{r, echo=FALSE}
# Takes a long time to run so here are some pre-determined results:
# save(results, file='HM_GLM_coefs.Rdata')
load('HM_GLM_coefs.Rdata')
results
```

---

Compare these to the simulation parameters:

```{r echo=FALSE}
cat('int: ', int, '\n')
cat('age_eff: ', paste(age_eff, collapse=', '), '\n')
cat('pop_eff: ', paste(pop_eff, collapse=', '), '\n')
cat('sex_eff: ', paste(sex_eff, collapse=', '), '\n')
cat('se1: ', se1, '\n')
cat('se2: ', se2, '\n')
cat('sp1: ', sp1, '\n')
cat('sp2: ', sp2, '\n')
```


## Exercise

Instructions here




## Summary

Take-away points

```{r cleanup, include=FALSE}
unlink('glmtemplate.bug')
unlink('basic_hw.bug')
unlink('automodel.bug')
```
